---
title: "King Count Tract Level Maps"
author: "Brian Bartley"
date: "4/14/2022"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(cluster)
library(viridis)
library(leaflet)
library(tmap)
library(sf)
library(factoextra)
demographics <- read_rds("King County Tract Level Demographics.rds")
#writing a little code to get rid of lang data
langs <- c("l_spanish",
           "l_russian",
           "l_korean",
           "l_chinese",
           "l_vietnamese",
           "l_other_lang")
demographics <- demographics %>% select (-langs)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(23994)
vars <- demographics %>% select(-GEOID, -tract, -county, -state, -geometry) %>% colnames()
to_cluster <- demographics %>% select(-county, -tract, -state, -geometry)
row.names(to_cluster) <- to_cluster$GEOID
to_cluster$GEOID <- NULL
distance_matrix <- cluster::daisy(demographics[,vars])
res.pam <- cluster::pam(x=distance_matrix, k = 5, cluster.only = F)
demographics$cluster <- as.double(res.pam$clustering)
```

# Clustering demographic data
```{r, echo=FALSE, message=FALSE, warning=FALSE}
demographics$cluster <- factor(demographics$cluster,
                                  levels=c(1,3,5,2,4),
                                  labels=c("low income",
                                           "low middle income",
                                           "middle",
                                           "high middle income",
                                           "high income"),
                                  ordered=T)
tmap_mode("plot")
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = "cluster",
          palette = "Set1",
          title = "relative income",
          legend.hist = TRUE) +
  tm_layout(title = "Cluster by\nDemographic\nCharachteristic",
            frame = FALSE,
            legend.outside = TRUE,
            bg.color = "grey70",
            legend.hist.width = .8) +
    tm_scale_bar(position = c("left", "bottom"))
```

# Dynamic Maps

## Demographic clusters
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#demographics %>% colnames()
tmap_mode("view")
x <- "cluster"
y <- "demographic cluster"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## Income
```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- "medinc"
y <- "median income"
z <- " quintile"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## Poverty
Percent of people living within 200% of the federal poverty line.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#demographics %>% colnames()
x <- "p_pov"
y <- "% below 2X poverty"
z <- " quintile"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## Disabilities 
Percent of the people living with a federally reconginzed disability.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#demographics %>% colnames()
x <- "p_ability"
y <- "% disabiliity"
z <- " quintile"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## Age
Percent of people 65 and up.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#demographics %>% colnames()
x <- "p_65_up"
y <- "% 65 and up"
z <- " quintile"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## Limited English proficiency 
Percent of people who speak english less than well.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#demographics %>% colnames()
x <- "p_e_less"
y <- "% limited Eng."
z <- " quintile"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## Native population
Percent of people who are Indigenous to the area or Native American.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# demographics %>% colnames()
x <- "amercian_Indian_alone"
y <- "% Native Amerian"
z <- " quintile"
demographics %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```

## People of Color
Percent of people who do not identify as White Non-Hispanic.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- "POC"
y <- "% non-white"
z <- " quintile"
demographics %>% mutate(POC = (1-white_alone)) %>%
  st_sf() %>% tm_shape() + 
  tm_polygons(col = x,
              style = "quantile",
          palette = "Set1",
          title = paste(y,z),
          alpha = .3,
          popup.vars = c("tract", x))+
  tm_basemap(leaflet::providers$OpenStreetMap)
```



